---
title: "Tau-U power analysis"
author: "Jürgen Wilbert"
date: "`r Sys.Date()`"
format:
  html:
    df-print: kable
    output-file: index.html
    toc: true
    toc-location: left
    self-contained: true
    code-fold: true
comments: 
  hypothesis:
    theme: clean
    openSidebar: false
  giscus: 
    repo: jazznbass/supplement-tau-u-power
execute: 
  freeze: true
  warning: false
  error: false
---

```{r}
#| include: false
#| label: setup

# when you run into problems with the scplot/mcscan github version, please install the package included in this repository
#if (!("scplot" %in% installed.packages())) 
#  devtools::install_github("jazznbass/scaledic")
if (!("mcscan" %in% installed.packages())) 
  devtools::install_github("jazznbass/mcscan")
if (!("scplot" %in% installed.packages())) 
  devtools::install_github("jazznbass/scplot")


library(mcscan)
library(scplot)

# if necessary, install the following cran packages ahead
library(scan)


tau_u_a_B <- function(x) {
    res <- tau_u(x, method = "parker", tau_method = "a", meta_method = "none")$table[[1]]
    list(
      p = res[which(row.names(res) == "A vs. B"), which(names(res) == "p")],
      es = res[which(row.names(res) == "A vs. B"), which(names(res) == "Tau")]
    )
}

tau_u_a_B_ta <- function(x) {
    res <- tau_u(x, method = "parker", tau_method = "a", meta_method = "none")$table[[1]]
    list(
      p = res[which(row.names(res) == "A vs. B - Trend A"), which(names(res) == "p")],
      es = res[which(row.names(res) == "A vs. B - Trend A"), which(names(res) == "Tau")]
    )
}

tau_u_a_B_ta_tb <- function(x) {
    res <- tau_u(x, method = "parker", tau_method = "a", meta_method = "none")$table[[1]]
    list(
      p = res[which(row.names(res) == "A vs. B + Trend B - Trend A"), which(names(res) == "p")],
      es = res[which(row.names(res) == "A vs. B + Trend B - Trend A"), which(names(res) == "Tau")]
    )
  }


tau_u_adj <- function(x) {
    res <- corrected_tau(x, continuity = FALSE, repeated = FALSE)
    list(p = res$p, es = res$tau)
}


evaluation <- function(x) {

    p <- unlist(lapply(x, function(y) {
      if (y$es <= 0) y$p <- 1
      y$p
    }))
    es <- unlist(lapply(x, function(y) y$es))
    c(p = mean(p <= 0.05, na.rm = TRUE) * 100, es = mean(es, na.rm = TRUE))
}


create_mc <- function(filename, 
                      design, 
                      iterations, 
                      force_renew = FALSE,
                      eval, ...) {
  
  #force_renew <- TRUE
  
  if (file.exists(filename) && !force_renew) {
    out <- readRDS(filename)
  } else {
    out <- mcstudy(
      design = design,
      iterations = iterations,
      methods = list(
        "Tau-U A vs. B" = tau_u_a_B,
        "Tau-U - trendA" = tau_u_a_B_ta, 
        "Tau-U - trendA + trendB" = tau_u_a_B_ta_tb, 
        "Tau-U adjusted" = tau_u_adj
      ),
      eval_function = evaluation,
      rf_arguments = list(round = 0),
      ...
    )
    saveRDS(out, filename)
  }
  out
}

cat_parameters <- function(iterations, design) {
  cat("**Iterations:**  \n")
  cat(paste(names(iterations), "=", iterations, collapse = "  \n"), "  \n")
  cat("**Design:**  \n")
  cat(paste(names(design), "=", design, collapse = "  \n"), "  \n")
}

```

# Research question

Q: How strong does a data trend influence the effect-size estimation and power of various tau-u analyzing methods?


# Length of A and B phase

```{r}
design <- design_template(
  n = n_sims,
  start_value = 50,
  s = 10,
  level = 1 * statistic,
  trend = rnorm(n_sims, 0, abs(0.1 / (A_length + B_length))),
  slope = rnorm(n_sims, 0, abs(0.1 / B_length)),
  phase_design = list(A = A_length, B = B_length),
  distribution = "gaussian"
)

iterations <- iterations_template(
  A_length = c(3, 5, 7, 9, 11, 13, 15, 17, 19),
  B_length = c(5, 7, 10, 15, 25, 50),
  statistic = c(0, 1),
  n_sims = 5000
)

out <- create_mc("phaselength.rds", 
                 design, iterations, 
                 force_renew = FALSE)
```


::: panel-tabset


## Example plots

```{r}
ex_design <- design(
  n = 3,
  start_value = 50,
  s = 10,
  level = 1,
  trend = rnorm(3, 0, abs(0.1 / c(8, 26, 69))),
  slope = rnorm(3, 0, abs(0.1 / c(5, 15, 50))),
  phase_design = list(A = c(3,11,19), B = c(5,15,50)),
  distribution = "gaussian"
)

ex <- random_scdf(ex_design,seed = 124)

ex_fn <- function(x) {
  scplot(x) %>% 
    set_yaxis(limits = c(40,70)) %>%
    add_theme("basic")
}


ex_fn(ex[1])
ex_fn(ex[2])
ex_fn(ex[3])


```


## Power plot

```{r}
mcplot(out, template = "power")

ggsave(filename = "figure-phaselength.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)
```

## Power table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html", digits = 0)
```

## Tau plot

```{r}
mcplot(out,var_y = "es")
ggsave(filename = "figure-phaselength-tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)
```

## Tau table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html", var_y = "es")
```

:::

# Intervention strength

```{r}
design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect * statistic,
    trend = rnorm(n_sims, 0, 0.1 / length),
    slope = rnorm(n_sims, 0, 0.1 / (length * 2/3)),
    phase_design = list(A = length * 1/3, B = length * 2/3),
    distribution = "gaussian"
)

iterations <- iterations_template(
  level_effect = c(0.25, 0.5, 0.75, 1, 1.25, 1.5),
  length = c(9, 12, 15, 21, 27, 33, 39, 45),
  statistic = c(0,1),
  n_sims = 5000
)

out <- create_mc("interventioneffect-length.rds", 
                 design, iterations,
                 force_renew = FALSE)
```

::: panel-tabset

## Example plots

```{r}
ex_design <- design(
  n = 3,
  start_value = 50,
  s = 10,
  level = c(0.25, 1, 1.5),
  trend = rnorm(3, 0, abs(0.1 / c(9, 21, 50))),
  slope = rnorm(3, 0, abs(0.1 / c(6, 14, 45))),
  phase_design = list(A = c(3,7,15), B = c(6,14,45)),
  distribution = "gaussian"
)

ex <- random_scdf(ex_design,seed = 124)

ex_fn <- function(x) {
  scplot(x) %>% 
    set_yaxis(limits = c(40,80)) %>%
    add_theme("basic")
}


ex_fn(ex[1])
ex_fn(ex[2])
ex_fn(ex[3])
```

## power plot

```{r}
mcplot(out, var_x = 2, var_shape = 1, template = "power")

ggsave(filename = "figure-interventioneffect.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)
```

## power table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html", digits = 0)
```


## Tau plot

```{r}
mcplot(out, var_y = "es", var_x = 2, var_shape = 1)

ggsave(filename = "figure-interventioneffect-tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)

```

## Tau table

```{r}
#| column: page

mctable(out, var_y = "es", wider = "statistic", format = "html")
```


:::

# Strength of trend effect and level-effect

```{r}
design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect * statistic,
    trend = trend_effect / length,
    slope = rnorm(n_sims, 0, abs(0.1/ (length* 2/3))),
    phase_design = list(A = length * 1/3, B = length * 2/3 ),
    distribution = "gaussian"
)

iterations <- iterations_template(
  trend_effect = c(-0.5, -0.25, 0, +0.25, +0.5),
  level_effect = c(0.25, 0.5, 0.75, 1, 1.25, 1.5),
  statistic = c(0, 1),
  length = 15,
  n_sims = 5000
)

out <- create_mc("trend-level.rds", 
                 design, iterations,
                 force_renew = FALSE)
```

::: panel-tabset

## Example plots

```{r}
ex_design <- design(
  n = 3,
  start_value = 50,
  s = 10,
  level = c(0.25, 0.5, 1),
  trend = c(-0.5/15, 0, 0.5/15),
  slope = rnorm(3, 0, abs(0.1 / 10)),
  phase_design = list(A = 5, B = 10),
  distribution = "gaussian"
)

ex <- random_scdf(ex_design,seed = 124)

ex_fn <- function(x) {
  scplot(x) %>% 
    set_yaxis(limits = c(40, 80)) %>%
    add_theme("basic")
}


ex_fn(ex[1])
ex_fn(ex[2])
ex_fn(ex[3])
```

## Power plot

```{r}
mcplot(out, var_y = "p", var_x = 1, var_shape = 2, template = "power")

ggsave(filename = "figure-trend-level.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)
```

## Power table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html", digits = 0)
```


## Tau plot

```{r}
mcplot(out, var_y = "es", var_x = 1, var_shape = 2)

ggsave(filename = "figure-trend-level-tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)
```

## Tau table

```{r}
mctable(out, var_y = "es", wider = "statistic", format = "html")
```


:::

# Strength of trend effect and meassurement times

```{r}
design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect * statistic,
    trend = trend_effect / length,
    slope = rnorm(n_sims, 0, 0.1 / (length * 2/3)),
    phase_design = list(A = length * 1/3, B = length * 2/3),
    distribution = "gaussian"
)

iterations <- list(
  trend_effect = c(-0.5, -0.25, 0, +0.25, +0.5),
  length = c(9, 12, 15, 21, 27, 33),
  statistic = c(0,1),
  level_effect = 1,
  n_sims = 5000
)

out <- create_mc("trend-length.rds", 
                 design, iterations,
                 force_renew = FALSE)
```


::: panel-tabset

## Example plots

```{r}
ex_design <- design(
  n = 3,
  start_value = 50,
  s = 10,
  level = 1,
  trend = c(-0.5/9, 0, 0.5/27),
  slope = rnorm(3, 0, abs(0.1 / c(6,10,28))),
  phase_design = list(A = c(3,5,9), B = c(6,10,18)),
  distribution = "gaussian"
)

ex <- random_scdf(ex_design,seed = 124)

ex_fn <- function(x) {
  scplot(x) %>% 
    set_yaxis(limits = c(40, 80)) %>%
    add_theme("basic")
}


ex_fn(ex[1])
ex_fn(ex[2])
ex_fn(ex[3])
```


## Power plot

```{r}
mcplot(out, var_x = 1, var_shape = 2, template = "power")

ggsave(filename = "figure-trend-length.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)
```

## Power table

```{r}
mctable(out, var_y = "p", wider = "statistic", format = "html", digits = 0)
```


## Tau plot

```{r}
mcplot(out, var_y = "es", var_x = 1, var_shape = 2)

ggsave(filename = "figure-trend-length-tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)
```

## Tau table

```{r}
#| column: page

mctable(out, var_y = "es", wider = "statistic", format = "html")
```

:::


# Strength of trend effect and meassurement times (equal phase length)

```{r}
design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect * statistic,
    trend = trend_effect / length,
    slope = rnorm(n_sims, 0, 0.1 / (length * 2/3)),
    phase_design = list(A = length * 1/2, B = length * 1/2),
    distribution = "gaussian"
)

iterations <- list(
  trend_effect = c(-0.5, -0.25, 0, +0.25, +0.5),
  length = c(8, 12, 16, 22, 28, 34),
  statistic = c(0,1),
  level_effect = 1,
  n_sims = 5000
)

out <- create_mc("trend-length-2.rds", 
                 design, iterations,
                 force_renew = FALSE)
```

::: panel-tabset


## Power plot

```{r}
mcplot(out, var_x = 1, var_shape = 2, template = "power")

ggsave(filename = "figure-trend-length-2.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)
```

## Power table

```{r}
mctable(out, var_y = "p", wider = "statistic", format = "html", digits = 0)
```


## Tau plot

```{r}
mcplot(out, var_y = "es", var_x = 1, var_shape = 2)

ggsave(filename = "figure-trend-length-2-tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)
```

## Tau table

```{r}
#| column: page

mctable(out, var_y = "es", wider = "statistic", format = "html")
```

:::
