---
title: "Tau-U power analysis"
author: "Jürgen Wilbert"
date: "`r Sys.Date()`"
format:
  html:
    df-print: kable
    output-file: index.html
    toc: true
    toc-location: left
    self-contained: true
    code-fold: true
comments:
  giscus:
    repo: jazznbass/supplement-tau-u-power
execute: 
  freeze: true
  warning: false
  error: false
---

```{r}
#| include: false
#| label: setup

# when you run into problems with the scplot/mcscan github version, please install the package included in this repository
#if (!("scplot" %in% installed.packages())) 
#  devtools::install_github("jazznbass/scaledic")
if (!("mcscan" %in% installed.packages())) 
  devtools::install_github("jazznbass/mcscan")

library(mcscan)

# if necessary, install the following cran packages ahead
library(scan)


create_mc <- function(filename, 
                      design, 
                      iterations, 
                      force_renew = FALSE,
                      eval, ...) {
  
  #force_renew <- TRUE
  
  #eval_function <- get_mcfn("pos_perc_sig")
  
  if (eval == "p") eval_function <- get_mcfn("pos_perc_sig")
  if (eval == "es") eval_function <- get_mcfn("mean_es")
  
  if (file.exists(filename) && !force_renew) {
    out <- readRDS(filename)
  } else {
    out <- mcstudy(
      design = design,
      iterations = iterations,
      methods = list(
        "Tau-U A vs. B" = get_mcfn("tau_u_AB"),
        "Tau-U trendA" = get_mcfn("tau_u_trendA"), 
        "Tau-U trendA + trendB" = get_mcfn("tau_u_trendA_trendB"), 
        "Tau-U adjusted" = get_mcfn("tau_u_base")
      ),
      eval_function = eval_function,
      ...
    )
    saveRDS(out, filename)
  }
  out
}

cat_parameters <- function(iterations, design) {
  cat("**Iterations:**  \n")
  cat(paste(names(iterations), "=", iterations, collapse = "  \n"), "  \n")
  cat("**Design:**  \n")
  cat(paste(names(design), "=", design, collapse = "  \n"), "  \n")
}

```

# Research question

Q: How strong does a data trend influence the effect-size estimation and power of various tau-u analyzing methods?


# Length of A and B phase

::: panel-tabset

## Power plot

```{r}

design <- design_template(
  n = n_sims,
  start_value = 50,
  s = 10,
  level = 1 * statistic,
  trend = rnorm(n_sims, 0, abs(0.1 / (A_length + B_length))),
  slope = rnorm(n_sims, 0, abs(0.1 / B_length)),
  phase_design = list(A = A_length, B = B_length),
  distribution = "gaussian"
)

iterations <- iterations_template(
  A_length = seq(3, 20, by = 2),
  B_length = c(10, 15, 20, 30, 40, 50),
  statistic = c(0, 1),
  n_sims = 1000
)

out <- create_mc("phaselength.rds", 
                 design, iterations, 
                 eval = "p",
                 force_renew = FALSE)

mcplot(out, template = "power")

ggsave(filename = "figure-phaselength.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)
```

## Power table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html", digits = 0)
```

## Tau plot

```{r}
out <- create_mc("phaselength-tau.rds", 
                 design, iterations,
                 force_renew = FALSE,
                 eval = "es")

mcplot(out)
ggsave(filename = "figure-phaselength-tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)

```

## Tau table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html")
```

:::

# Intervention strength

::: panel-tabset

## power plot

```{r}

design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect * statistic,
    trend = rnorm(n_sims, 0, 0.1 / length),
    slope = rnorm(n_sims, 0, 0.1 / (length * 2/3)),
    phase_design = list(A = length * 1/3, B = length * 2/3),
    distribution = "gaussian"
)

iterations <- iterations_template(
  level_effect = c(0.25, 0.5, 0.75, 1, 1.25, 1.5),
  length = c(15, 21, 27, 33, 39, 45),
  statistic = c(0,1),
  n_sims = 1000
)

out <- create_mc("interventioneffect-length.rds", 
                 design, iterations, eval = "p",
                 force_renew = FALSE)

mcplot(out, var_x = 2, var_shape = 1, template = "power")

ggsave(filename = "figure-interventioneffect.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)

```

## power table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html", digits = 0)
```


## Tau plot

```{r}
out <- create_mc("interventioneffect-length-tau.rds", 
                 design, iterations, eval = "es",
                 force_renew = FALSE)

mcplot(out, var_x = 2, var_shape = 1)

ggsave(filename = "figure-interventioneffect-tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)



```

## Tau table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html")
```


:::

# Strength of trend effect and level-effect

::: panel-tabset

## Power plot

```{r}
design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect * statistic,
    trend = trend_effect / length,
    slope = rnorm(n_sims, 0, abs(0.1/ (length* 2/3))),
    phase_design = list(A = length * 1/3, B = length * 2/3 ),
    distribution = "gaussian"
)

iterations <- iterations_template(
  trend_effect = c(-0.5, -0.25, 0, +0.25, +0.5),
  level_effect = c(0.25, 0.5, 0.75, 1, 1.25, 1.5),
  statistic = c(0, 1),
  length = 15,
  n_sims = 1000
)

out <- create_mc("trend-level.rds", 
                 design, iterations, eval = "p",
                 force_renew = FALSE)

mcplot(out, var_x = 1, var_shape = 2, template = "power")

ggsave(filename = "figure-trend-level.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)


```

## Power table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html", digits = 0)
```


## Tau plot

```{r}
out <- create_mc("trend-level-tau.rds", 
                 design, iterations, eval = "es",
                 force_renew = FALSE)

mcplot(out, var_x = 1, var_shape = 2)

ggsave(filename = "figure-trend-level-tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)


```

## Tau table

```{r}
mctable(out, wider = "statistic", format = "html")
```


:::

# Strength of trend effect and meassurement times

::: panel-tabset

## Power plot

```{r}
design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect * statistic,
    trend = trend_effect / length,
    slope = rnorm(n_sims, 0, 0.1 / (length * 2/3)),
    phase_design = list(A = length * 1/3, B = length * 2/3),
    distribution = "gaussian"
)

iterations <- list(
  trend_effect = c(-0.5, -0.25, 0, +0.25, +0.5),
  length = c(15, 21, 27, 33, 39, 45),
  statistic = c(0,1),
  level_effect = 1,
  n_sims = 1000
)

out <- create_mc("trend-length.rds", 
                 design, iterations, eval = "p",
                 force_renew = FALSE)

mcplot(out, var_x = 1, var_shape = 2, template = "power")

ggsave(filename = "figure-trend-length.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)



```

## Power table

```{r}
mctable(out, wider = "statistic", format = "html", digits = 0)
```


## Tau plot

```{r}
out <- create_mc("trend-length-tau.rds", 
                 design, iterations, eval = "es",
                 force_renew = FALSE)

mcplot(out, var_x = 1, var_shape = 2)

ggsave(filename = "figure-trend-length-tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)


```

## Tau table

```{r}
#| column: page

mctable(out, wider = "statistic", format = "html")
```

:::
