---
title: "Tau-U power analysis"
author: "Jürgen Wilbert"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include = FALSE}
# when you run into problems with the scplot github version, please install the package included in this repository
if (!("scplot" %in% installed.packages())) 
  devtools::install_github("jazznbass/scaledic")

# if necessary, install the following cran packages ahead
library(scan)
library(scplot)
library(tidyverse)
library(knitr)
library(kableExtra)
library(mcscan)

#source("plot_mc_function.R")
#source("table_mc_function.R")
#source("mc_analysis_functions.R")
#source("mc_function.R")
#source("mc_power_test.R")

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

methods_p <- list(
  "Tau-U A vs. B" = get_mcfn("tau_u_AB"),
  "Tau-U trendA" = get_mcfn("tau_u_trendA"), 
  "Tau-U trendA + trendB" = get_mcfn("tau_u_trendA_trendB"), 
  "Tau-U adjusted" = get_mcfn("tau_u_base")
)

methods_es <- list(
  "Tau-U A vs. B" = get_mcfn("tau_u_AB_es"),
  "Tau-U trendA" = get_mcfn("tau_u_trendA_es"), 
  "Tau-U trendA + trendB" = get_mcfn("tau_u_trendA_trendB_es"), 
  "Tau-U adjusted" = get_mcfn("tau_u_base_es")
)


create_mc <- function(filename, 
                      design, 
                      iterations, 
                      force_renew = FALSE,
                      eval, ...) {
  
  if (eval == "p") {
    eval_function <- get_mcfn("perc_sig")
    methods <- methods_p
  }
  if (eval == "es") {
    eval_function <- get_mcfn("mean")
    methods <- methods_es
  }
  
  if (file.exists(filename) && !force_renew) {
    out <- readRDS(filename)
  } else {
    out <- mcstudy(
      design = design,
      iterations = iterations,
      methods = methods,
      eval_function = eval_function,
      ...
    )
    saveRDS(out, filename)
  }
  out
}

cat_parameters <- function(iterations, design) {
  cat("**Iterations:**  \n")
  cat(paste(names(iterations), "=", iterations, collapse = "  \n"), "  \n")
  cat("**Design:**  \n")
  cat(paste(names(design), "=", design, collapse = "  \n"), "  \n")
}

```


# Research question

Q: How strong does a data trend influence the effect-size estimation and power of various tau-u analyzing methods?


# Length of A and B phase

## Percentage significant

```{r}

design <- design_template(
  n = n_sims,
  start_value = 50,
  s = 10,
  level = level_effect,
  trend = rnorm(n_sims, 0, abs(0.1 / (A_length + B_length))),
  slope = rnorm(n_sims, 0, abs(0.1 / B_length)),
  phase_design = list(A = A_length, B = B_length),
  distribution = "gaussian"
)

iterations <- iterations_template(
  A_length = seq(3, 20, by = 2),
  B_length = c(10, 15, 20, 30, 40, 50),
  level_effect = c(0, 1),
  n_sims = 10
)

out <- create_mc("phaselength.rds", 
                 design, iterations, 
                 eval = "p",
                 force_renew = TRUE)

mcplot(out, template = "power")

ggsave(filename = "figure_phaselength.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)

#mctable(out)

```

## Mean Tau

```{r}
out <- create_mc("phaselength_tau.rds", 
                 design, iterations,
                 force_renew = TRUE,
                 eval = "es")

mcplot(out)
```

## Parameters

```{r results='asis'}
cat_parameters(iterations, design)
```

# Intervention strength

## Percentage significant


```{r}

design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect,
    trend = rnorm(n_sims, 0, 0.1 / length),
    slope = rnorm(n_sims, 0, 0.1 / (length * 2/3)),
    phase_design = list(A = length * 1/3, B = length * 2/3),
    distribution = "gaussian"
)

iterations <- iterations_template(
  level_effect = c(0, 0.25, 0.5, 0.75, 1, 1.25, 1.5),
  length = c(15, 21, 27, 33, 39, 45),
  n_sims = 10
)

out <- create_mc("interventioneffect_length.rds", 
                 design, iterations, eval = "p",
                 force_renew = TRUE)

mcplot(out, var_x = 2, var_shape = 1, template = "power")

ggsave(filename = "figure_interventioneffect.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)

```

## Mean Tau-U

```{r}
out <- create_mc("interventioneffect_length_tau.rds", 
                 design, iterations, eval = "es",
                 force_renew = TRUE)

mcplot(out, var_x = 2, var_shape = 1)

ggsave(filename = "figure_interventioneffect_tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1)



```


## Parameters

```{r results='asis'}
cat_parameters(iterations, design)
```


# Strength of trend effect and measurement-times

## Percentage significant

```{r}
design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect,
    trend = trend_effect / length,
    slope = rnorm(n_sims, 0, abs(0.1/ (length* 2/3))),
    phase_design = list(A = length * 1/3, B = length * 2/3 ),
    distribution = "gaussian"
)

iterations <- iterations_template(
  trend_effect = c(-0.5, -0.25, 0, +0.25, +0.5),
  length = c(15, 30, 45),
  level_effect = c(0, 0.5, 1, 1.5),#c(0, 0.25, 0.5, 0.75, 1, 1.25, 1.5),
  n_sims = 100
)

out <- create_mc("trend.rds", 
                 design, iterations, eval = "p",
                 force_renew = TRUE)

mcplot(out, var_x = 1, var_shape = 2)

ggsave(filename = "figure_trend.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)


```

## Mean Tau-U

```{r}
out <- create_mc("trend_length_tau.rds", 
                 design, iterations, eval = "es",
                 force_renew = TRUE)

mcplot(out, var_x = 1, var_shape = 2)

ggsave(filename = "figure_trend_tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)

```


## Parameters

```{r results='asis'}
cat_parameters(iterations, design)
```


# Strength of trend effect and level-effect

## Percentage significant

```{r}
design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = level_effect,
    trend = trend_effect / 15,
    slope = rnorm(n_sims, 0, abs(0.1/ (10))),
    phase_design = list(A = 5, B = 10 ),
    distribution = "gaussian"
)

iterations <- list(
  trend_effect = c(-0.5, -0.25, 0, +0.25, +0.5),
  level_effect = c(0.25, 0.5, 0.75, 1, 1.25, 1.5),
  n_sims = 20
)

out <- create_mc("trend_level.rds", 
                 design, iterations, eval = "p",
                 force_renew = FALSE, alpha_test = TRUE)

plot_mc(out, var_x = 2, var_shape = 1)

ggsave(filename = "figure_trend_level.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)


table_mc(out, reference_category = 0, first = 1, second = 2)
table_mc(out, reference_category = 1, first = 1, second = 2)

```

## Mean Tau-U

```{r}
design <- design_template(
    n = n_sims,
    start_value = 50,
    s = 10,
    level = 1,
    trend = trend_effect / length,
    slope = rnorm(n_sims, 0, abs(0.1/ (length* 2/3))),
    phase_design = list(A = length * 1/3, B = length * 2/3 ),
    distribution = "gaussian"
)

iterations <- list(
  trend_effect = c(-0.5, -0.25, 0, +0.25, +0.5),
  length = c(15, 21, 27, 33, 39, 45),
  n_sims = 200
)

out <- create_mc("trend_tau.rds", 
                 design, iterations, eval = "es",
                 force_renew = FALSE, alpha_test = TRUE)

plot_mc(out, ylim = NA, marks = 0, ylab = "Tau",
        var_x = 2, var_shape = 1,
        statistic_label = c("No effect", "Effect"))

ggsave(filename = "figure_trend_tau.png", width = 18, height = 18,  device='png', units = "cm", dpi=300, scale = 1.1)


table_mc_tau(out, reference_category = 0, first = 1, second = 2)
table_mc_tau(out, reference_category = 1, first = 1, second = 2)

```


## Parameters

```{r results='asis'}
cat_parameters(iterations, design)
```

